---
title: "EDA (Explanatory Data Analysis) of the dataset heights"
authors: Marit Rygg Fredheim & Namra Shahid
format: html
editor: visual
bibliography: datascience2.bib
csl: apa-no-ampersand.csl
output:
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
# load packages
    library(tidyverse)
    library(modelr)
    library(grid)
    library(ggplotify)
    library(vtable)
    library(gridExtra)
    library(ggplot2)
    library(magrittr)
    library(dplyr)
    library(expm)
    library(broom)
    library(knitr)
```

```{r, include = FALSE}
heights <- modelr::heights
```

## Abstract

How we can use EDA (Explanatory Data Analysis).

## Introduction

In this essay are we going to study the dataset of heights in the modelr package.
We will do an EDA (Explanatory Data Analysis) of the dataset.
To solve this assignment we are going to use the program RStudio.

## Literature Review

In this Literature review we are going to discuss physical height and income.
More explained, we are going to discuss the connection between physical gender, education and income.

There are many researchers who investigate or research on the connection between physical gender, education and income.
[@averettEconomicRealityBeauty1996].
Most researchers investigate income, marital status, and hourly pay differentials by body mass (kg/m2).
According to researchers differences in economic status by body mass for women is increasing markedly when we use an old weight data/measurement.
The same applies also if we restrict the sample to persons who were single and childless when the old weight was sampled. 

A research from The university of Chicago Press Journal show that stature, status, self-esteem, social dominance and discrimination is based on the association between height, earnings and labour market outcomes [@caseStatureStatusHeight2008].

A study from Deaton shows that the height of an adult depends on genetics and by intake of food, like nutrition and the importance of balanced food [@deatonHeightHealthDevelopment2007].
This study also shows that low nutrition quality can cause non survival of children which can cause short height or in worst case death.

One study shows that taller workers receive a wage earning premium just by increasing/developing strength, size, stamina and enhancing labor productivity [@schickHeightHumanCapital2015] .
Most studies show that height and disturbance.
Here many researchers react to this, and Persico et al. (2004) says that this is developed and increased at a young age.
Many scientists come up with a theory that when it comes to productivity it's important to mention emotional stability and boldness because these two are important factors.

There is much social awareness on being overweight.
For example, one important thing to mention is that Americans face much social and psychological awareness when it comes to being overweight [@mankiwOptimalTaxationHeight].
In this theory we see that height is one of the factors to get independent and earn more.
But at the same time height is one of many characteristics.
There are many other characteristics that matter more than just the height for a person to earn money.

Researchers show that taller persons are smarter than shorter ones [@pinskerEveryInchTaller2015].
Taller people do better at a young age, compared to smaller ones.
There are many scientists out there who will disagree with this theory, but there is some evidence that shows that taller persons are smarter and earn more.

Many studies show that marriage has a large impact on work [@reedMarriagePremiumCompensating].
The important fact here to come up with is that marital status relies upon the consumption of the family.
This study shows that men are more likely to marry compared to women.

Recent research shows us that men´ wages rise more rapidly than expected prior to marriage [@killewaldNewEvidenceCausal2017].
The main focus of this research is that married men experience or go through wage advantages, compared with never-married men and divorced men.

Not only marriage but other studies show that stature and favorable labor market outcomes has also been extensively documented [@lundborgHeightPremiumEarnings2009].
Height is a very important factor, when it comes to being productive and earning money.

Many experts believe that height is a significant factor in professional success [@wangWhatCreatingHeight2020].
However, why do higher wages for being tall occur?
Yet why is a greater salary associated with greater height?
More intriguingly, why is it observed in labor markets that lack a height factor thought to be critical to productivity, for example, sedentary white-collar jobs in industrialized countries?
This question has inspired many researchers to interpret the observed height premium as proof of employment discrimination by employers that target shorter individuals.
However, this interpretation is now disputed on the basis of a purported association between height and improved employment outcomes; the alleged association may be the result of unmeasurable factors that affect one's employment performance.
For instance, scientists have found that height is correlated to both cognition and non-cognitive abilities; these abilities play a large role in one's career success.

The result of @wangWhatCreatingHeight2020 shows us that height is indeed associated with higher salary income.
Association between income and one's physical appearance (like height, stature, BMI, or facial appearance) has inspired a large number of studies that aim to discover the causal mechanisms.
This document continues the discussion on this topic and uses recent information from China to demonstrate that without taking into account dependence, height is correlated with higher income.
Using the unique genetic data associated with prosperous individuals, we then explore and demonstrate that genetic cognition, genetic risk of depression, various personality traits, as well as other confounders that are associated with height are likely to be the actual cause of the observed height premium.
Through the unique advantageous individual genetic data, we then investigate and display that genetic cognitive ability, genetic risk of depression, various personality traits, as well as other confounding variables that are correlated with height are likely to be the true causes of the observed height advantage.

There is one more study that shows that height has a big impact on one person's salary [@traversYourHeightHas2020].
This study also talks about Wang´s research in China.
There is also an even more interesting fact.
Researchers used genetic testing to come up with the effect of height on income.

## EDA

In this study we are using the dataset heights.
The purpose of this dataset is to compare how seven different variables affect each other.
This dataset consist of the variables income, height, weight, age, marital, sex, education and aftq (percentile score on Armed Forces Qualification Test).
Our consern will be to describe the variables with the help of tables and graphics and inn addition start to investigate how the seven variables influence income.
We will start to present a summary of the data from the dataset.\

```{r}
st(heights)
```

### Marital

```{r}
ggplot(data = heights) + 
  geom_bar(mapping = aes(x = marital, fill = marital))
```

***Graph 1.0***

What we can see from this graph 1.0 is the categories of marital state.
If we take a look at this graph, we see that there are more married individuals.
When we look at married graph in this survey we see that there are approximately 3800 married individuals.
To make this graph more specific are we going to make the marital categories to only "married" and "not married".

```{r}
heights <- heights %>%
  mutate(
    married = fct_collapse(
      .f = marital,
      married = "married",
      other_level = "not-married"
    )
  ) 
```

```{r}
heights %>% 
  mutate(married = if_else(marital == 'married', TRUE, FALSE)) %>% 
  select( -marital) %>% 
  vtable::st(group = 'married')
```

```{r}
heights %>% 
  mutate(
    married = if_else(marital == 'married', TRUE, FALSE)) %>% 
  filter(income > 0) %>% 
  ggplot(mapping = aes(x = married, y = income)) +
  facet_wrap(~sex) + 
  geom_boxplot()
```

***Graph 1.1***

What we can see in this graph is two categories married and not married male and female and how much income they have.
We can see category married on x-axis and category income in y-axis.
What we can say by looking at this graph is that men who are married earn more/ have more income in comparison with non married men.
And when we look at female, it looks like there is no such big difference between married and non married females, when it comes to how they earn/ have in income

### Education

The next thing we are going to look into is education.

```{r}
ggplot(data = heights) +
  geom_bar(mapping = aes(x = education), na.rm = TRUE)
```

We see from the table that we also have a lot of categories in this variable.
We have 20 different categories and we want to divide this into smaller subcategories.
There are a peak at 12, 14 and 16 years.
Because of that we will divide the variable into these four subcategories: not high-school (0-12 years), not trade collage (12-14 years), not collage (14-16 years) and collage + (16-21 years).
The following graphs are also divided into men and female with blue and pink colour.

```{r}
heights %>% 
  mutate(
    edu_fac = cut(education, 
                  breaks = c(0, 12, 14, 16, 21), 
                  labels = c("NotHS", "NotTC", "NotC", "C+"),
                  right = FALSE) 
  ) %>% 
  filter(!is.na(edu_fac) & income > 0) %>%
  ggplot(mapping = aes(x = income, fill = sex, colour = sex)) +
  geom_density(alpha = 0.2, na.rm = TRUE) + 
  facet_wrap(~edu_fac)
```

[Litt om kva vi ser av grafane... Flatere kurve på dei som har collage + utdanning, kan sjå ut som det er fleire i denne grafen som har høgare inntekt... Dess lågare inntekt dess høgare er kurva i starten av grafen. (Kan sjå ut til at dess lengre og høgare utdanning, dess høgare inntekt?)]{.underline}

### Sex/Height/Income/Education

[Her vil eg eigentle ha education på høgre sida derfor står det education i overskrifta.]{.underline}

```{r}
heights <- heights %>% 
  group_by(sex) %>% 
  mutate(
    heightIntSex = cut_number(height, n = 4, labels = c("very low", "low", "medium", "high"))
  )
```

```{r}
heights %>% 
  filter(income > 0 & income < 300000) %>%
  ggplot(mapping = aes(x = height, y = income)) +
  geom_point(mapping = aes(colour = factor(age)), size = 0.7, alpha = 0.3, na.rm = TRUE) + 
  geom_smooth(method = "lm", colour = "black", lwd = 0.5, na.rm = TRUE) + 
  facet_grid(heightIntSex ~ sex) + 
  theme(legend.position = "bottom", strip.text.y = element_text(angle = 0))
```

[Desse grafane viser]{.underline}... høgde og utdanning i forhold til inntekt.
Grafen viser også korles alder fordelar seg.

### Weight & Aftq

[Til slutt manglar vi berre å sjå på vekt og afqt (percentile score on Armed Forces Qualification Test). Desse skal vi sjå på ved å bruke histogram.]{.underline}

```{r}
hist_weight <- ggplotify::as.ggplot(~hist(heights$weight, breaks = 20))
hist_afqt <- ggplotify::as.ggplot(~hist(heights$afqt, breaks = 20))

gridExtra::grid.arrange(hist_weight, hist_afqt, ncol = 2)
```

[Forklare grafane]{.underline}

## Discussion

From EDA we can say that the connection between physical height, gender, education and income is that income from what we have researched in this assignment depends on the factors such as height.
As we have discussed in the Literature review, many researchers shows us that height determine income.
Most research are in favor of that height have much to say about how much income one individual have.
In the Literature review we have discussed some researchers point of view on height dependence on income.

## Conclusion

We can here conclude by saying that height determine income and also other data determine income.
Other data such as: weight, age, education and afqt, as we have seen in our statistics under EDA also in some ways determine income.
As we have seen different researchers point of view on height determination on income, can we conclude by saying that height can in many contexts determine income.

## References

::: {#refs}
:::

## Tools used

```{r}
print(citation(), bibtex = TRUE)
```

```{r}
print(citation("modelr"), bibtex = TRUE)
```
