---
title: "Assignment2"
author: Marit Rygg Fredheim
author: Namra Shahid
format: html
editor: visual
---

## Abstract

How we can use EDA (Explanatory Data Analysis).

## Introduction

In this essay are we going to study datasets of heights in the modelr package.
We will in this assignment do an EDA (Explanatory Data Analysis) of the dataset heights in the modelr package.
We are going to use program R to solve this assignment. 

## Literature Review

What is known about the connection between physical height, gender, education and income.

In this Literature review we are going to discuss physical height and income.
More explained, we are going to discuss here the connection between physical gender, education and income. 

##   EDA

```{r}
heights <- modelr::heights
```

```{r}
heights <- heights %>%
mutate(
married = fct_collapse(
.f = marital,
married = "married",
other_level = "not-married"
)
)
```

## Discussion

## Conslusion

## Refrences

## Tools used

```{r}
print(citation(), bibtex = TRUE)
```

```{r}
print(citation("modelr"), bibtex = TRUE)
```
